|

### 2023-04-27 manuscript/manuscript_1.0.Rmd
Changes to the reference list

Associative learning is a taxonomically widespread cognitive 
mechanism that allows individuals to associate rewards with environmental stimuli 
and thus behave adaptively 
[@heyes_Simple_2012;@macphail_Brain_1982; @staddon_Adaptive_2016; 
@behrens_Associative_2008].
Theory has shown that natural selection favours such associations 
in complex environments where conditions are difficult to predict 
[@dridi_Environmental_2016]. Associative learning is a flexible cognitive 
mechanism whose underpinnings show interspecific variation 
[@enquist_Power_2016; @quinones_Reinforcement_2019]. Associative learning 
is not a single mechanisms, but rather a set of cognitive structures 
and processes that can vary in their scope and complexity. Presumably, 
these structures and processes have been modified 
by natural selection and could provide new explanations for behavioural 
variation. Despite all this, associative learning is 
not often included in the narrative of evolutionary explanations of 
behavioural patterns [@fawcett_Exposing_2013; @kamil_Optimal_1983; 
@mcauliffe_Psychology_2015]. Computational models of evolution 
can play an important role to overcome the lack of integration 
between learning and evolution. Reinforcement learning theory 
encompasses a series of computational methods inspired 
on the psychological and neurological mechanisms of associative 
learning [@sutton_Reinforcement_2018]. This set of algorithms allows the 
implementation of biologically realistic problems, 
capturing the essence of learning processes 
[@frankenhuis_Enriching_2018; @quinones_Reinforcement_2019]. 
Furthermore, these algorithms can be embedded in evolutionary simulations to 
generate theoretical predictions of the effect that learning can 
have in behavioural evolution [@leimar_Learning_2019;@leimar_Effects_2022].

@@

The decision of whether to escalate a fight against an interacting 
partner can depend on the size of the partner's badge, 
the dependency is determined by the focal's experiences 
through a learning process. We implement learning using 
the actor-critic approach from reinforcement learning (RL) theory [@sutton_Reinforcement_2018;@quinones_Reinforcement_2019; @leimar_Learning_2019].
Individuals estimate the reward (pay-off) expected from  interacting 
with partners of different badge sizes (the critic in RL terminology). 
After each interaction they update the estimate of reward proportionally 
to the difference between their current estimate and the observed 
reward (prediction error $\delta$) and to the speed of learning 
($\mathrm{A}$). Furthermore, individuals express different probabilities of 
retreating/attacking depending on the badge size of their opponent 
(the actor in RL). They update the probability of 
retreating/attacking  up or down depending on whether retreating/attacking
leads to an increase in the reward estimation. 
So, if a focal individual decides to escalate a fight against 
an individual with small badge and this leads to an increase in 
the reward estimation, then the focal individual will increase the probability 
of escalating fights with individuals of small badges in the future. 
Given that badge size is a real number between 0 and 1, there are infinitely 
many badge sizes. Thus, the reward estimation, as well as the probability 
of retreating/attacking, must be generalized across different values. 
To implement generalization we use the linear function approximation 
method based on radial basis functions [@sutton_Reinforcement_2018]. 
Specifically, we pick $c$ feature centres, which are evenly spaced values 
along the badge size interval ([0,1]) where the updates are focused. 
For the sake of simplicity we keep the location of these feature centres 
constant and  the same for all individuals, the are stored in the 
vector $\symbf{b}$ . Each of these feature centres 
is associated with a weight for reward estimation and tendency to play 
retreat in a given interaction. The reward estimation and the tendency 
to retreat are calculated (in every interaction) as the sum of the weights 
associated with each feature centre, weighted by the response 
triggered by the feature (Fig. \ref{fig:model-struc} B, dots 
represent the feature weights). The response of each feature centre 
diminishes as a Gaussian function with the distance between the 
feature centre and the badge size of 
the partner (Fig. \ref{fig:model-struc} B, grey line). Formally, the 
reward estimation $\hat{R}$ when the focal individual ($i$) faces 
individual $j$ is given by,
\
\begin{equation}
\hat{R}_{ij} =  \sum_{z=1}^{c} x_z e^{(-\frac{|B_i-b_z|}{2\theta^2})} 
\end{equation}

@@

```{r model-struc,echo=FALSE,warning=FALSE,fig.width=8,fig.align='center',fig.height=7,fig.show='hold',fig.cap= "Model of communication in the context of aggressive interactions. In A, the reaction norm determines the badge size. Red shows an informative reaction norm, while the blue shows a uninformative reaction norm. B individuals perceiving the signal have a behavioral reaction norm that determines their probability of retreating (black line). The reaction norm arises from generalizing the information from the feature weights (black dots). Feature weaights are values associated with a particular badge size and are updated as individuals interact with each other. The generalization of the information associated with each feature weight is represented by the grey line and its axis, which shows the response triggered by the fourth feature diminishes as the value evaluated is further from the feature center. The learning process moves the feature weights (black dots) up or down depending on whether that leads to an increase in the estimated reward. In C and D, effect of learning on the receiver strategy. Receivers in C face signallers with uninformative reaction norms (blue line in A). The uninformative reaction norm yields badge size of 0.5, thus the behavioural reaction norm in C is only updated around that value.D Receivers face signallers with informative reaction norms (red line in A). Accordingly, receivers develop a threashold-like reaction norm where the decision to retreat increases with the badge size of the interacting partner. Colour scale in C and D indicates the quality of the individual. "}

@@

Associative learning is a powerful mechanism to learn about the world, about 
social partners and about one's abilities relative to the social context. 
By associating cues and signals with fitness relevant outcomes individuals
collect information that allows them to improve their reproductive potential. 
This is very obvious in classical examples such as when animals learn to avoid 
food that makes them sick, or use environmental cues to find food. 
More recent ideas have highlighted the potential role that associations 
can have in a social contexts such as: hierarchy formation 
[@leimar_Reproductive_2022;@leimar_Effects_2022],
and cooperation [@leimar_Learning_2019;@dridi_Learning_2018]. In this 
theoretical examples individuals use various sources of reward to make
adaptive choices in a given social context. Here we extend this logic to the 
evolution of a communication system mediated by a quality signal. 
Previous evolutionary models of badges of status assumed that individuals 
responded using a behavioural reaction norm, where the opponent's badge 
and individual's own quality determined the behaviour [@botero_Evolution_2010]. 
However, there isn't a clear mechanism to justify the assumption that 
individuals inherently know their own quality, particularly relative to their peers.
In our model, individuals not only learn about the quality signal, but also 
learn about their own quality relative to others. This can be seen throughout
our simulations in the responses developed by individuals of different quality. 
Because individuals learn about their own quality, 
even in the absence of an honest signal, they are able to make adaptive choices. 
Thus the learning process we have modelled and the response it mediates 
has fitness relevant consequences even in the absence of an honest signal.

>>>>

Associative learning is a taxonomically widespread cognitive 
mechanism that allows individuals to associate rewards with environmental stimuli 
and thus behave adaptively 
[@heyes_Simple_2012;@macphail_Brain_1982; @staddon_Adaptive_2016; 
@behrens_Associative_2008].
Theory has shown that natural selection favours such associations 
in complex environments where conditions are difficult to predict 
[@dridi_Environmental_2016]. Associative learning is a flexible cognitive 
mechanism whose underpinnings show interspecific variation 
[@enquist_Power_2016; @quinones_Reinforcement_2020]. Associative learning 
is not a single mechanisms, but rather a set of cognitive structures 
and processes that can vary in their scope and complexity. Presumably, 
these structures and processes have been modified 
by natural selection and could provide new explanations for behavioural 
variation. Despite all this, associative learning is 
not often included in the narrative of evolutionary explanations of 
behavioural patterns [@fawcett_Exposing_2013; @kamil_Optimal_1983; 
@mcauliffe_Psychology_2015]. Computational models of evolution 
can play an important role to overcome the lack of integration 
between learning and evolution. Reinforcement learning theory 
encompasses a series of computational methods inspired 
on the psychological and neurological mechanisms of associative 
learning [@sutton_Reinforcement_2018]. This set of algorithms allows the 
implementation of biologically realistic problems, 
capturing the essence of learning processes 
[@frankenhuis_Enriching_2018; @quinones_Reinforcement_2019]. 
Furthermore, these algorithms can be embedded in evolutionary simulations to 
generate theoretical predictions of the effect that learning can 
have in behavioural evolution [@leimar_Learning_2019;@leimar_Effects_2022].

@@

The decision of whether to escalate a fight against an interacting 
partner can depend on the size of the partner's badge, 
the dependency is determined by the focal's experiences 
through a learning process. We implement learning using 
the actor-critic approach from reinforcement learning (RL) theory [@sutton_Reinforcement_2018;@quinones_Reinforcement_2020; @leimar_Learning_2019;
@leimar_Evolution_2021].
Individuals estimate the reward (pay-off) expected from  interacting 
with partners of different badge sizes (the critic in RL terminology). 
After each interaction they update the estimate of reward proportionally 
to the difference between their current estimate and the observed 
reward (prediction error $\delta$) and to the speed of learning 
($\mathrm{A}$). Furthermore, individuals express different probabilities of 
retreating/attacking depending on the badge size of their opponent 
(the actor in RL). They update the probability of 
retreating/attacking  up or down depending on whether retreating/attacking
leads to an increase in the reward estimation. 
So, if a focal individual decides to escalate a fight against 
an individual with small badge and this leads to an increase in 
the reward estimation, then the focal individual will increase the probability 
of escalating fights with individuals of small badges in the future. 
Given that badge size is a real number between 0 and 1, there are infinitely 
many badge sizes. Thus, the reward estimation, as well as the probability 
of retreating/attacking, must be generalized across different values. 
To implement generalization we use the linear function approximation 
method based on radial basis functions [@sutton_Reinforcement_2018]. 
Specifically, we pick $c$ feature centres, which are evenly spaced values 
along the badge size interval ([0,1]) where the updates are focused. 
For the sake of simplicity we keep the location of these feature centres 
constant and  the same for all individuals, the are stored in the 
vector $\symbf{b}$ . Each of these feature centres 
is associated with a weight for reward estimation and tendency to play 
retreat in a given interaction. The reward estimation and the tendency 
to retreat are calculated (in every interaction) as the sum of the weights 
associated with each feature centre, weighted by the response 
triggered by the feature (Fig. \ref{fig:model-struc} B, dots 
represent the feature weights). The response of each feature centre 
diminishes as a Gaussian function with the distance between the 
feature centre and the badge size of 
the partner (Fig. \ref{fig:model-struc} B, grey line). Formally, the 
reward estimation $\hat{R}$ when the focal individual ($i$) faces 
individual $j$ is given by,
\
\begin{equation}
\hat{R}_{ij} =  \sum_{z=1}^{c} x_z e^{(-\frac{|B_i-b_z|}{2\theta^2})} 
\end{equation}

@@

```{r model-struc,echo=FALSE,warning=FALSE,fig.width=8,fig.align='center',fig.height=7,fig.show='hold',fig.cap= "Model of communication in the context of aggressive interactions. In A, the reaction norm determines the bib (badge) size, black band on birds chest. Red shows an informative reaction norm, while the blue shows a uninformative reaction norm. B individuals perceiving the signal have a behavioral reaction norm that determines their probability of retreating (black line). The reaction norm arises from generalizing the information from the feature weights (black dots). Feature weaights are values associated with a particular badge size and are updated as individuals interact with each other. The generalization of the information associated with each feature weight is represented by the grey line and its axis, which shows the response triggered by the fourth feature diminishes as the value evaluated is further from the feature center. The learning process moves the feature weights (black dots) up or down depending on whether that leads to an increase in the estimated reward. In C and D, effect of learning on the receiver strategy. Receivers in C face signallers with uninformative reaction norms (blue line in A). The uninformative reaction norm yields badge size of 0.5, thus the behavioural reaction norm in C is only updated around that value.D Receivers face signallers with informative reaction norms (red line in A). Accordingly, receivers develop a threashold-like reaction norm where the decision to retreat increases with the badge size of the interacting partner. Colour scale in C and D indicates the quality of the individual. "}

@@

Associative learning is a powerful mechanism to learn about the world, about 
social partners and about one's abilities relative to the social context. 
By associating cues and signals with fitness relevant outcomes individuals
collect information that allows them to improve their reproductive potential. 
This is very obvious in classical examples such as when animals learn to avoid 
food that makes them sick, or use environmental cues to find food. 
More recent ideas have highlighted the potential role that associations 
can have in a social contexts such as: hierarchy formation 
[@leimar_Evolution_2021;@leimar_Reproductive_2022;@leimar_Effects_2022],
and cooperation [@leimar_Learning_2019;@dridi_Learning_2018]. In this 
theoretical examples individuals use various sources of reward to make
adaptive choices in a given social context. Here we extend this logic to the 
evolution of a communication system mediated by a quality signal. 
Previous evolutionary models of badges of status assumed that individuals 
responded using a behavioural reaction norm, where the opponent's badge 
and individual's own quality determined the behaviour [@botero_Evolution_2010]. 
However, there isn't a clear mechanism to justify the assumption that 
individuals inherently know their own quality, particularly relative to their peers.
In our model, individuals not only learn about the quality signal, but also 
learn about their own quality relative to others. This can be seen throughout
our simulations in the responses developed by individuals of different quality. 
Because individuals learn about their own quality, 
even in the absence of an honest signal, they are able to make adaptive choices. 
Thus the learning process we have modelled and the response it mediates 
has fitness relevant consequences even in the absence of an honest signal.
